\chapter{Terminology and Abbreviations}\label{app:terminology}

\begin{description}
\item[VR -] Virtual Reality.
\item[VE -] Virtual Environment.
\item[RDW -] Redirected Walking.
\item[ROT -] Reorientation Technique. These are used in redirected walking whenever the user is close to leaving the physical tracking space in order to reorient them away from it. 
\item[Reset Technique -] A type of ROT. These forcefully reorient the user away from physical walls. While they can break the subjective sense of presence, they are primarily used as fail-safes if other methods are insufficient.
\item[NPC -] Non-playable character. A term used in video games to classify a computer-controlled entity.
\item[HUD -] Heads-Up Display.
\item[GUI -] Graphical User Interface.
\item[S2C -] ''Steer To Center''. A redirection algorithm that applies gains in a manner that redirects the user towards the centre of the tracking space.
\item[S2O -] ''Steer To Orbit''. A redirection algorithm that applies gains in a manner that redirects the user to walk on the edge of a circle in the tracking space.
\item[AC2F -] ''Align Centre To Future''. A redirection algorithm that has been developed for this thesis. It focuses on redirecting a standing user so their future walking direction is aligned with the centre of the physical space.
\item[FoV -] Field of View.
\item[HMD -] Head Mounted Display. Also often known as a virtual reality headset.
\item[IDE -] Integrated Development Environment.
\item[2AFC -] Two-alternative Forced Choice.
\end{description}